---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here("~/projects/2024_mrg_and_arg_copy_from_saahith/"))

library(tidyverse)
library(Rsamtools)
library(ggpubr)
library(readr)
library(ggmap)
library(viridis)
library(RColorBrewer)



pg <- DBI::dbConnect(RPostgres::Postgres(),dbname = "glamr_data", host = "localhost", port = "5432", user = "glamr_admin", password = "glamr2023") 
```

```{r}
secrets <- yaml::read_yaml("secrets.yaml")
maps_api_key <- secrets$maps_api_key
```


## Idetifying Samples for mapping
Get sample metadata
```{r}
# Read in sample and study tables
# samples <- readxl::read_excel("data/Great_Lakes_Omics_Datasets.xlsx",sheet = "samples",guess_max = 3000) %>% 
#   mutate(date = lubridate::ymd(collection_date),
#          date = if_else(is.na(date), lubridate::ymd_hms(collection_date), date),
#          date = if_else(is.na(date), lubridate::ymd_hm(collection_date), date),
#          date_same_year = lubridate::`year<-`(date,2014),
#          year = lubridate::year(date)) %>% 
#   select(SampleID, StudyID, date, date_same_year, year, lat, lon, NOAA_Site , starts_with("size_frac"))


# Metadata exists for some samples that haven't had reads downloaded; don't consider these in subsequent steps
samples_mapped <- Sys.glob("data/metagenomic_reads/samp_*_fwd.fastq.gz") %>%
  data.frame(path = .) %>%
  unglue::unglue_unnest(path, "data/metagenomic_reads/{SampleID}_fwd.fastq.gz",remove = FALSE) %>%
  mutate(link_path = fs::link_path(path),
         data_exists = fs::file_exists(link_path)) %>%
  filter(data_exists) %>%
  pull(SampleID)

# Read in sample information from postgres database
samples <- tbl(pg,"glamr_samples") |>
  filter(sample_type == "metagenome",
         SampleID %in% local(samples_mapped)) %>% 
  collect()
```
Read-in Gene Info
```{r}
gene_info <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1fwBbxcnydPch3_zZQ_SgdU6M7SDZDaz6BVY8SM6x6sE/edit?pli=1#gid=0") 

id_count <- gene_info |> 
  group_by(gene_id) |>
  summarize(n_obvs = n())

```
Read-in Total Number of Reads Per Sample
```{r}
pg <- DBI::dbConnect(RPostgres::Postgres(),dbname = "glamr_data", host = "localhost", port = "5432", user = "glamr_admin", password = "glamr2023")


read_counts <- tbl(pg,"read_count") |>
  filter(sample %in% local(samples$SampleID),
         read_state == "decon_reads",
         direction == "fwd") |>
  collect()
```


Define Function for Proccessing.read_mapping 
```{r}
bam_path = "data/gene_mapping/samp_1497_mapped.bam"


bam_stats_w_nuc_info <- function(bam_path){

  bam <- Rsamtools::BamFile(file = bam_path,
                            index = paste0(bam_path,".bai"))
  
  # number of mapped_reads
  reads_pile <- idxstatsBam(bam) |>
    mutate(gene_id=str_remove(seqnames,":.*")) |>
    left_join(gene_info)
  
  
  
  
  # Can be used to get seq lengths quickly
  # bam_count <- scanBamHeader(bam)$targets %>% 
  #   data.frame(length = ., seq = labels(.))
  
  out <- Rsamtools::pileup(bam,pileupParam = PileupParam(distinguish_strands = FALSE)) %>% 
    mutate(sample_total_bases_mapped = sum(count)) %>% 
    group_by(seqnames) %>% 
    mutate(seq_length = max(pos),
           per_ref_total_mapped_bases = sum(count),
           ref_rel_abund = per_ref_total_mapped_bases / sample_total_bases_mapped) %>% 
    group_by(seqnames, pos, seq_length) %>% 
    arrange(seqnames, pos, desc(count)) %>% 
      mutate(non_primary_base_count = sum(count) - max(count),
             nuc_rank = row_number(),
             depth = sum(count),
             rel_abund = count/depth,
             prop_non_primary_base = sum(non_primary_base_count) / depth,
             ratio_to_primary = count / max(count),
             percent_covered_bases = (max(depth) > 0) / seq_length * 100) %>%
    #left_join(. %>% group_by(seq_length,.drop = FALSE) %>% summarise(percent_covered_bases = sum(min(depth) > 0) / seq_length * 100))
    group_by(seqnames) %>% 
      mutate(prop_non_primary_seq = sum(non_primary_base_count) / per_ref_total_mapped_bases,
             #percent_covered_bases = sum(depth > 0) / seq_length * 100,
             mean_depth = mean(depth),
             depth_var = var(depth),
             depth_sd = sd(depth),
             percent_id_to_consensus = 100 - prop_non_primary_seq * 100,
             bam_path = bam_path) %>% 
    left_join(reads_pile)
}
```
Define a Function for Calculating Normalized Read Counts (rpkm)
```{r}
calc_rpkm <- function(bam_path){
   bam <- Rsamtools::BamFile(file = bam_path,
                            index = paste0(bam_path,".bai"))
  
  # number of mapped_reads
  reads_pile <- idxstatsBam(bam) |>
    mutate(gene_id=seqnames, 
      ## gene_id=str_remove(seqnames,":.*"),
           SampleID=str_remove(bam_path,"data/gene_mapping/") |> 
           str_remove("_mapped.bam")) |>
    left_join(gene_info) |>
    left_join(read_counts |>
                select(SampleID="sample",sample_read_count="count")) |>
    mutate(per_million = sample_read_count/1000000,
           rpm = mapped/per_million,
           rpkm = rpm/(seqlength/1000)) # seqlength in kilobasepair
} 
```

Read-in read_mapping results
```{r}
bam_files <- Sys.glob("data/gene_mapping/*.bam")

bam_files_w_info <- bam_files %>% 
  data.frame(path = .) %>% 
  bind_cols(., unglue::unglue_data(.$path,"data/gene_mapping/{SampleID}_mapped.bam"))

mapped_sample_info <- samples %>% 
  filter(SampleID %in% bam_files_w_info$SampleID) 


all_bam_stats_w_seqs_bam <- map_df(bam_files_w_info$path, bam_stats_w_nuc_info) |>
  write_rds("data/read_mapping_summary.rds")

all_bam_stats_w_seqs_bam <- read_rds("data/read_mapping_summary.rds")
```
Reads-in the proccessed results (if restarting)


all_bam_stats_w_seqs_bam <- read_rds("data/read_mapping_summary.rds")

Creates Summary Table

summary.table <- all_bam_stats_w_seqs_bam |>
  select(seqnames,gene_id,gene,rel_abund,percent_covered_bases,mean_depth,percent_id_to_consensus,bam_path,seqlength,tax_id,species,lineage) |>
  distinct() %>%
  mutate(sample=str_remove(bam_path,"data/gene_mapping/") |> 
           str_remove("_mapped.bam"))  |>
  left
  


Calc. rpkm
```{r}
rpkm_results <- map_dfr(bam_files_w_info$path,calc_rpkm) 

write_rds(rpkm_results,"data/rpkms.rds") 

rpkm_results <- read_rds("data/rpkms.rds") 
```

Summarize rpkm results
```{r}
rpkm_sample_info <- rpkm_results |>
  left_join(mapped_sample_info) 

per_gene_rpkm_summary <- rpkm_sample_info |>
  group_by(gene,SampleID) |>
  summarize(rpkm=sum(rpkm)) |>
  left_join(mapped_sample_info) 
```

```{r}
GLAMR_datasets <- tbl(pg, "glamr_datasets") |>
  collect()

public_glmr_datasets <- GLAMR_datasets |>
  filter(private==FALSE)

public_metagenome_samples <- samples |>
  filter(SampleID %in% samples$SampleID, 
         SampleID %in% samples_mapped,
         StudyID %in% public_glmr_datasets$StudyID, 
         geo_loc_name=="Lake Erie") |>
  write_csv("data/for_pub_supplemental/filtered_samples.tsv")

```


Presence Plot of MRGs
```{r}
library(ggplot2)
library(dplyr)

metal_genes <- c("MerA", "MerP", "MerB", "MerC", "MerD", "MerE", "MerR", "MerT", "MerF", "ArsD", "PbrT", "PbrA")
metal_data <- per_gene_rpkm_summary %>%
  filter(gene %in% metal_genes) %>%
  mutate(year = as.numeric(format(date, "%Y"))) %>%  
  filter(year %in% c(2014, 2019))

metal_data$year <- factor(metal_data$year, levels = c(2014, 2019))

(p1 <- metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggplot(aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  theme_bw() +
  scale_y_continuous(limits = c(0, 0.01)) +  
  labs(title = "Presence of Metal Resistant Genes in the Microbiome", x = "Year", y = "RPKM")
)
  
print(p1)


ggsave("results/Presence_of_Metal_Resistant_Genes_in_the_Microbiome.png", plot = p1, width = 10, height = 6, dpi = 300)
ggsave("results/Presence_of_Metal_Resistant_Genes_in_the_Microbiome.pdf", plot = p1, width = 10, height = 6)
```




Creates Summary Plot for MRGs
```{r}
library(ggplot2)
library(dplyr)

# Assuming that the 'date' column is already a Date object and 'year' is a numeric column
metal_genes <- c("MerA", "MerP", "MerB", "MerC", "MerD", "MerE", "MerR", "MerT", "MerF", "ArsD", "PbrT", "PbrA")
metal_data <- per_gene_rpkm_summary %>%
  filter(gene %in% metal_genes) %>%
  mutate(year = as.numeric(format(date, "%Y")),
         year = as.factor(year)) %>%  # Ensure 'year' is a numeric column
  filter(year %in% c("2014", "2019"))

years_to_show <- c("2014", "2019")

# Plot 2: Individual Trends of All the Metal Resistant Genes
(p2 <- metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggplot(aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(method = "lm") +
  #scale_x_continuous(breaks = years_to_show, limits = range(years_to_show)) +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Individual Trends of All the Metal Resistant Genes", y = "RPKM", x = "Year")
)


# Save the second plot
ggsave("results/Individual_Trends_of_All_the_Metal_Resistant_Genes.png", plot = p2, width = 10, height = 6, dpi = 300)
ggsave("results/Individual_Trends_of_All_the_Metal_Resistant_Genes.pdf", plot = p2, width = 10, height = 6)

# Plot 3: Overall Trend of All the Metal Resistant Genes

data_2014 <- metal_data %>% 
  filter(SampleID %in% public_metagenome_samples$SampleID,
         year == "2014") %>% 
  pull(rpkm)
  
data_2019 <- metal_data %>% 
  filter(SampleID %in% public_metagenome_samples$SampleID,
         year == "2019") %>% 
  pull(rpkm)

# Check distribution of the data
data_2014 %>% log() %>% hist()

# Run t-test example
t.test(data_2014, data_2019)

# Run wilcox test example
wilcox.test(data_2014, data_2019)


(p3 <- metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggplot(aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "lm", color = "black") +
  #scale_x_continuous(breaks = years_to_show, limits = range(years_to_show)) +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Overall Trend of All the Metal Resistant Genes", y = "RPKM", x = "Year") 
  #theme(legend.position = "bottom", aspect.ratio = 1.0) # Set aspect ratio here
)

metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggboxplot(x = "year", y = "rpkm",outlier.shape = NA) +
  stat_compare_means(method = "wilcox",na.rm = TRUE,vjust = 0.8) +
  scale_y_log10() +
  ggbeeswarm::geom_quasirandom(aes(color = gene), alpha = 0.5) 

ggsave("results/Stats_Overall_Trend_of_All_the_Metal_Resistant_Genes.png", width = 10, height = 12, dpi = 300)
ggsave("results/Stats_Overall_Trend_of_All_the_Metal_Resistant_Genes.pdf", width = 10, height = 12)



metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggplot(aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "lm", color = "black") +
  #scale_x_continuous(breaks = years_to_show, limits = range(years_to_show)) +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Overall Trend of All the Metal Resistant Genes", y = "RPKM", x = "Year")
  #theme(legend.position = "bottom", aspect.ratio = 1.0) # Set aspect ratio here

# Original plot code remains the same

# Adjust the width of the saved plot image to half


# Plot: Annual Variations of the Metal Resistant Genes Microbiome
metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) |>
  filter(rpkm > 1e-5) %>%
  ggplot(aes(x = date_same_year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10() +
  facet_wrap(~year, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8), 
        axis.title.x = element_text(size = 10), 
        strip.text.x = element_text(size = 10)) + 
  labs(x = "Time of Year", 
       title = "Annual Variations of the Metal Resistant Genes Microbiome", 
       y = "RPKM")
ggsave("results/Annual_Variations_of_the_Metal_Resistant_Genes_Microbiome.png", plot = last_plot(), width = 14, height = 8, dpi = 300)
ggsave("results/Annual_Variations_of_the_Metal_Resistant_Genes_Microbiome.pdf", plot = last_plot(), width = 14, height = 8)


my_comparisons <- list(c("2014", "2019"))

metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID,
         rpkm > 0) %>%
  #mutate(gene_year = str_glue("{gene}-{year}")) %>% 
  ggboxplot(x = "year", y = "rpkm",outlier.shape = NA, color = "year") +
  stat_compare_means(comparisons = my_comparisons,method = "wilcox",na.rm = TRUE) +
  scale_y_log10() +
  scale_x_discrete(guide = guide_axis(angle = -35)) +
  facet_grid(~gene, scales = "free_x")
  ggbeeswarm::geom_quasirandom(aes(color = gene), alpha = 0.5) 

ggsave("results/BoxIndividual_Trends_of_All_MRGs.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
ggsave("results/BoxIndividual_Trends_of_All_MRGs.pdf", plot = last_plot(), width = 10, height = 6)

```



Individual/overall trend MRGs with STATS
```{r}
library(ggplot2)
library(ggpubr)
library(ggbeeswarm)

# Filter the data
filtered_data <- metal_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID)

# Create the plot and store it in a variable
(plot <- ggboxplot(filtered_data, x = "year", y = "rpkm", outlier.shape = NA) +
  stat_compare_means(method = "wilcox", na.rm = TRUE, vjust = 0.8) +
  scale_y_log10() +
  ggbeeswarm::geom_quasirandom(aes(color = gene), alpha = 0.5)
)

# Save the plot as PNG and PDF
ggsave("results/Stats_Overall_Trend_of_All_the_Metal_Resistant_Genes.png", plot = plot, width = 10, height = 12, dpi = 300)
ggsave("results/Stats_Overall_Trend_of_All_the_Metal_Resistant_Genes.pdf", plot = plot, width = 10, height = 12)

```



Creates Summary Plot for ARGs
```{r}
library(ggplot2)
library(dplyr)

other_genes <- c("AmpC", "ermA", "ermC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                 "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")
other_data <- per_gene_rpkm_summary %>%
  filter(gene %in% other_genes) %>%
  mutate(year = as.numeric(format(date, "%Y"))) %>%
  filter(year %in% c(2014, 2019))

other_data$year <- factor(other_data$year, levels = c(2014, 2019))

other_data %>%
 filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggplot(aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  theme_bw() +
  scale_y_continuous(limits = c(0, 0.01), labels = scales::comma) +  # Ensure the y-axis limits and labels are appropriate
  scale_x_discrete(name = "Year") +  # Changed from scale_y_log10 to scale_y_continuous
  labs(title = "Presence of Antibiotic Resistant Genes in the Microbiome", x = "Year", y = "RPKM") +
  theme(legend.position = "right")  # Adjust legend position if needed
ggsave("results/Presence_of_ARGs_in_the_Microbiome.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
ggsave("results/Presence_of_ARGs_in_the_Microbiome.pdf", plot = last_plot(), width = 8, height = 6)


other_data %>%
filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  ggplot(aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(method = "lm", aes(group = gene)) +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Individual Trends of All the Antibiotic Resistant Genes", y="RPKM", x="Year") +
  scale_x_discrete(expand = c(0.05, 0.05 )) + # Adjust the x-axis
  theme(legend.position = "bottom")

ggsave("results/Individual_Trends_of_All_ARGs.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
ggsave("results/Individual_Trends_of_All_ARGs.pdf", plot = last_plot(), width = 10, height = 6)


# with stats

my_comparisons <- list(c("2014", "2019"))

other_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID,
         rpkm > 0) %>%
  #mutate(gene_year = str_glue("{gene}-{year}")) %>% 
  ggboxplot(x = "year", y = "rpkm",outlier.shape = NA, color = "year") +
  stat_compare_means(comparisons = my_comparisons,method = "wilcox",na.rm = TRUE) +
  scale_y_log10() +
  scale_x_discrete(guide = guide_axis(angle = -35)) +
  facet_grid(~gene, scales = "free_x")
  ggbeeswarm::geom_quasirandom(aes(color = gene), alpha = 0.5) 
  
  
  # ggplot(aes(x = year, y = rpkm, color = gene)) +
  # geom_point() +
  # geom_smooth(method = "lm", aes(group = gene)) +
  # scale_y_log10() +
  # theme_bw() +
  # labs(title = "Individual Trends of All the Antibiotic Resistant Genes", y="RPKM", x="Year") +
  # scale_x_discrete(expand = c(0.05, 0.05 )) + # Adjust the x-axis
  # theme(legend.position = "bottom")

ggsave("results/BoxIndividual_Trends_of_All_ARGs.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
ggsave("results/BoxIndividual_Trends_of_All_ARGs.pdf", plot = last_plot(), width = 10, height = 6)


other_data %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) |>
  filter(rpkm > 1e-5) %>%
  filter(!is.na(year)) %>%
  ggplot(aes(x = date_same_year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10() +
  facet_wrap(~year, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8), # Rotate x labels for readability
        axis.title.x = element_text(size = 10), # Adjust x-axis title size if necessary
        strip.text.x = element_text(size = 10)) + # Adjust facet label size if necessary
  labs(x = "Time of Year", 
       title = "Annual Variations of the Antibiotic Resistant Genes Microbiome", 
       y = "RPKM")
ggsave("results/Annual_Variations_ARGs_Microbiome.png", plot = last_plot(), width = 14, height = 8, dpi = 300)
ggsave("results/Annual_Variations_ARGs_Microbiome.pdf", plot = last_plot(), width = 14, height = 8)



# Assuming that 'date' is a column of class Date or POSIXct.
other_data <- other_data %>%
  mutate(year = as.numeric(format(date, "%Y"))) %>%
  filter(year %in% c(2014, 2019)) %>%
  mutate(year = factor(year)) %>%
  filter(SampleID %in% public_metagenome_samples$SampleID)

# Generate the plot with only 2014 and 2019 data points
(p2 <- ggplot(other_data, aes(x = year, y = rpkm, color = gene)) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "lm", color = "black") +  # Single line of best fit
  scale_y_log10() +
  scale_x_discrete(expand = c(0.1, 0.1)) +  # Adjusted to reduce the gap on the x-axis
  theme_bw() +
  labs(title = "Overall Trend of All the Antibiotic Resistant Genes", y = "RPKM", x = "Year") +
  theme(
    legend.position = "bottom",  # Move legend to bottom to avoid overlapping
    plot.title = element_text(size = 10),  # Adjust the plot title font size
    legend.text = element_text(size = 5),  # Adjust the legend text font size
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 8)
  )
)


# Save the plot with adjusted dimensions
ggsave("results/Overall_Trend_of_All_ARGs.png", plot = p2, width = 4, height = 6, dpi = 300)
ggsave("results/Overall_Trend_of_All_ARGs.pdf", plot = p2, width = 4, height = 6)


```





Individual/overall trend ARGs with STATS

```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggbeeswarm)

# Define the genes to include
other_genes <- c("AmpC", "ermA", "ermC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                 "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")

# Filter and mutate the data
other_data <- per_gene_rpkm_summary %>%
  filter(gene %in% other_genes) %>%
  mutate(year = as.numeric(format(date, "%Y"))) %>%
  filter(year %in% c(2014, 2019)) %>%
  filter(SampleID %in% public_metagenome_samples$SampleID)

# Convert year to a factor
other_data$year <- factor(other_data$year, levels = c(2014, 2019))

# Run Wilcoxon test
data_2014 <- other_data %>% filter(year == "2014") %>% pull(rpkm)
data_2019 <- other_data %>% filter(year == "2019") %>% pull(rpkm)

wilcox_test_result <- wilcox.test(data_2014, data_2019)

# Print the Wilcoxon test result
print(wilcox_test_result)

# Create the plot and store it in a variable
(plot <- ggboxplot(other_data, x = "year", y = "rpkm", outlier.shape = NA) +
  stat_compare_means(method = "wilcox", na.rm = TRUE, vjust = 0.8) +
  scale_y_log10() +
  ggbeeswarm::geom_quasirandom(aes(color = gene), alpha = 0.5) +
  theme_bw() +
  labs(x = "Year", y = "RPKM")
)
  
# Save the plot as PNG and PDF
ggsave("results/FStats_Overall_Trend_of_All_ARGs.png", plot = plot, width = 10, height = 12, dpi = 300)
ggsave("results/FStats_Overall_Trend_of_All_ARGs.pdf", plot = plot, width = 10, height = 12)

```


Individual/overall trend MRGs with STATS

```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggbeeswarm)


```



Box Plot rpkm for MRGs
```{r}

# per_gene_rpkm_summary |>
#   filter(SampleID %in% public_metagenome_samples$SampleID) |>
#  # filter(rpkm>0) |>
#   filter(!is.na(year)) |>
#   filter(gene %in% c("MerA", "MerP","MerB","MerC","MerD","MerE","MerR", "MerT","MerF","ArsD","PbrT","PbrA")) |>
# ggplot(aes(x=gene,y=rpkm,color=gene)) + 
#   geom_boxplot() +
#   
#    scale_y_log10() + 
#   # facet_wrap(~year,scales="free_y") +
#   theme_bw() +
#   labs(title = "Box Plot of MRGs", x="Gene", y="RPKM") 





# MRG boxplot with ANOVA results

per_gene_rpkm_summary |>
  filter(SampleID %in% public_metagenome_samples$SampleID) |>
 # filter(rpkm>0) |>
  filter(!is.na(year)) |>
  filter(gene %in% c("MerA", "MerP","MerB","MerC","MerD","MerE","MerR", "MerT","MerF","ArsD","PbrT","PbrA")) |>
  ggboxplot(x = "gene", y = "rpkm", color = "gene") +
  scale_y_log10() +
  stat_compare_means(method = "kruskal") + # are there differences between the groups (genes)?
  stat_compare_means(method = "wilcox", comparisons = list(c("ArsD", "MerA"))) # if you want to add individual comparisons


# ggplot(aes(x=gene,y=rpkm,color=gene)) + 
#   geom_boxplot() +
#    scale_y_log10() + 
#   # facet_wrap(~year,scales="free_y") +
#   theme_bw() +
#   labs(title = "Box Plot of MRGs", x="Gene", y="RPKM") 


ggsave("results/Box_plot_of_MRGS.pdf",width=5,height=3,scale=1.5)
ggsave("results/Box_plot_of_MRGS.png",width=5,height=3,scale=1.5)

```




Box Plot of MRGs with STATS
```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)

# Filter the data
filtered_data <- per_gene_rpkm_summary |>
  filter(SampleID %in% public_metagenome_samples$SampleID) |>
  filter(!is.na(year)) |>
  filter(gene %in% c("MerA", "MerP", "MerB", "MerC", "MerD", "MerE", "MerR", "MerT", "MerF", "ArsD", "PbrT", "PbrA"))

# Perform the Kruskal-Wallis test
kruskal_test <- kruskal.test(rpkm ~ gene, data = filtered_data)
p_value <- kruskal_test$p.value
p_value_text <- ifelse(p_value < 2.2e-16, "< 2.2e-16", format(p_value, digits = 3))

# Create the plot and store it in a variable
(plot <- ggboxplot(filtered_data, x = "gene", y = "rpkm", color = "gene") +
  scale_y_log10() +
  
  annotate("text", x = 7, y = 1e-1, label = paste("Kruskal-Wallis test, p =", p_value_text), hjust = 0.5) + # Add custom annotation
  theme_bw() +
  labs(title = "Box Plot of MRGs", x="Gene", y="RPKM")
)
  
# Save the plot as PNG and PDF
ggsave("results/Stats_Box_Plot_of_MRGs.png", plot = plot, width = 15, height = 12, dpi = 300)
ggsave("results/Stats_Box_Plot_of_MRGs.pdf", plot = plot, width = 15, height = 12)

```





Box Plot rpkm for ARGs
```{r}
genes_to_include <- c("AmpC", "ermA", "ermC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                      "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")

ggplot(per_gene_rpkm_summary |>
   filter(SampleID %in% public_metagenome_samples$SampleID) |>

          # filter(rpkm>0) |>
          filter(!is.na(year)) |>
          filter(gene %in% genes_to_include), 
       aes(x=gene, y=rpkm, color=gene)) + 
  geom_boxplot() +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), 
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_discrete(guide = guide_axis(angle = 90)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(x="Gene", y="RPKM", title = "Box Plot of ARGs")
ggsave("results/Box_plot_of_ARGS.pdf",width=8,height=5,scale=1.5)
ggsave("results/Box_plot_of_ARGS.png",width=5,height=3,scale=1.5)
```






Box Plot for ARGs with STATS
```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)

# List of genes to include
genes_to_include <- c("AmpC", "ermA", "ermC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                      "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")

# Filter the data
filtered_data <- per_gene_rpkm_summary |>
  filter(SampleID %in% public_metagenome_samples$SampleID) |>
  filter(!is.na(year)) |>
  filter(gene %in% genes_to_include)

# Perform the Kruskal-Wallis test
kruskal_test <- kruskal.test(rpkm ~ gene, data = filtered_data)
p_value <- kruskal_test$p.value
p_value_text <- ifelse(p_value < 2.2e-16, "< 2.2e-16", format(p_value, digits = 3))

# Create the plot
(plot <- ggplot(filtered_data, aes(x=gene, y=rpkm, color=gene)) + 
  geom_boxplot() +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), 
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_discrete(guide = guide_axis(angle = 90)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(x="Gene", y="RPKM", title = "Box Plot of ARGs") +
  annotate("text", x = 10, y = 1e-1, label = paste("Kruskal-Wallis test, p =", p_value_text), hjust = 0.5) # Add custom annotation
)

# Save the plot as PNG and PDF
ggsave("results/Stats_Box_plot_of_ARGS.pdf", plot = plot, width = 8, height = 5, scale = 1.5)
ggsave("results/Stats_Box_plot_of_ARGS.png", plot = plot, width = 5, height = 3, scale = 1.5)

```








Bar Plot of Positive Samples
```{r}
per_gene_rpkm_summary |>
  filter(rpkm>1e-05) |>
  filter(!is.na(year)) |>
  select(gene,SampleID) |>
  distinct() |>
  group_by(gene) |>
  summarise(n_samples=n()) |>
ggplot(aes(x=gene,y=n_samples)) + 
  geom_bar(stat="identity") +
  
  # facet_wrap(~year,scales="free_y") +
  theme_bw() +
  labs(x="Gene",y="Number of Samples",title="Number of Samples Containing Metal Resistance Genes")

ggsave("results/Metal_resistance_gene_samples.pdf",width=5,height=3,scale=1.5)
ggsave("results/Metal_resistance_gene_samples.png",width=5,height=3,scale=1.5)
```
Map for MerD
```{r}
register_google(maps_api_key) 

ggmap(get_googlemap(center = c(lon = -83.326801, lat = 41.778739), zoom = 10, maptype = "terrain")) +
  geom_point(data = per_gene_rpkm_summary |> 
    filter(SampleID %in% public_metagenome_samples$SampleID) |>

               filter(gene=="MerD"), 
             aes(lon,lat, color = gene, size = rpkm), alpha = 0.5, inherit.aes = FALSE) +
    # geom_jitter(data = microcystis_abund, aes(lon,lat, color = org, size = percent_abundance, group = year), alpha = 0.5, inherit.aes = FALSE,width = 0.1, height = 0.05) +
  theme_bw() +
  #scale_color_viridis_c() +
  labs(x = "Longitude", y = "Latitude", color = "Gene", size = "RPKM", title = "Map of MerD Presence") +
  coord_quickmap(xlim = c(-83.5, -82.9), ylim = c(41.6, 42.089)) 
  

ggsave("results/MerD_map.pdf",width=5,height = 5,scale=1.5) 
ggsave("results/MerD_map.png",width=5,height = 5,scale=1.5) 
```
Map for all MRGs
```{r}
register_google(maps_api_key) 

# Define the metal-resistant genes
metal_resistant_genes <- c("MerA", "MerP", "MerB", "MerC", "MerD", "MerE", "MerR", "MerT", "MerF", "ArsD", "PbrT", "PbrA")

# Get the Google map
google_map <- get_googlemap(center = c(lon = -83.326801, lat = 41.778739), zoom = 10, maptype = "terrain")

# Prepare the data for plotting
gene_data <- per_gene_rpkm_summary %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  filter(gene %in% metal_resistant_genes) %>%
  left_join(gene_info) %>%
  select(gene, lon, lat, rpkm, type_name) %>%
  distinct()

avg_gene_abundance <- gene_data %>% 
  group_by(gene) %>% 
  filter(!is.na(rpkm)) %>% 
  summarise(mean_abund = mean(rpkm)) %>% 
  arrange(desc(mean_abund)) %>% 
  mutate(rank = row_number())

# Create a color palette that contrasts bright and dark colors
colors_df <- data.frame(color_hex = c("#3182BD", "#31A354", "#E6550D", "#FD8D3C", "#FDD0A2", "#74C476", "#6A51A3", "#B8E186","#9ECAE1", "#C7E9C0", "#FEE6CE", "#FDCDAC", "#FFF7BC", "#D9F0D3", "#DADAEB", "#A1D99B")) %>% 
  mutate(rank = row_number()) %>% 
  left_join(avg_gene_abundance)

gene_data_w_colors <- gene_data %>% 
  left_join(colors_df %>% select(gene, color_hex))


# Plot the map with points for metal-resistant genes
ggmap(google_map) +
  geom_point(data = gene_data_w_colors, 
             aes(lon, lat, color = color_hex, size = rpkm), alpha = 0.5, inherit.aes = FALSE) +
  scale_size_continuous(range = c(.2, 5)) + # Adjust size scale for better visibility
  scale_color_identity(guide = "legend",
                       labels = colors_df$gene,
                       breaks = colors_df$color_hex) +
  guides(color = guide_legend(override.aes = list(alpha = 1)),
         size = guide_legend(override.aes = list(alpha = 1), title = "RPKM")) +
  theme_bw() +
  theme(legend.position = "right",
        legend.key.width = unit(1, "cm"),
        legend.key.height = unit(0.43, "cm"),
        legend.text = element_text(size = 8)) +
  labs(x = "Longitude", y = "Latitude", color = "Gene", size = "RPKM") +
  coord_quickmap(xlim = c(-83.5, -82.9), ylim = c(41.6, 42.089)) +
  facet_grid(~type_name)

# Save the map to a file
ggsave("results/All_MRGs_map.pdf", width = 12, height = 7, scale = 1) 
ggsave("results/All_MRGs_map.png", width = 11, height = 6, scale = 1) 

ggsave("results/publication_figures/All_MRGs_map.pdf", width = 12, height = 7, scale = 1) 
ggsave("results/publication_figures/All_MRGs_map.png", width = 11, height = 6, scale = 1) 
```

Map for all ARGs
```{r}
library(ggmap)
library(ggplot2)
library(dplyr)
library(viridis)
library(RColorBrewer)

register_google(maps_api_key) 

# Define the antibiotic-resistant genes
antibiotic_resistant_genes <- c("AmpC", "ermA", "ermC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                                "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")

google_map <- get_googlemap(center = c(lon = -83.326801, lat = 41.778739), zoom = 10, maptype = "terrain")

# Create a custom color palette
color_palette <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Set2"), "#A6CEE3", "#B2DF8A", "#FB9A99", "#FDBF6F", "#CAB2D6")

# Ensure that the palette is the right length
color_palette <- color_palette[1:length(antibiotic_resistant_genes)]


# Create the plot with the new color palette
ggmap(google_map) +
  geom_point(data = per_gene_rpkm_summary %>% 
              filter(SampleID %in% public_metagenome_samples$SampleID) %>%
              filter(gene %in% antibiotic_resistant_genes) %>% 
              left_join(gene_info), 
             aes(lon, lat, color = gene, size = rpkm), alpha = 0.5, inherit.aes = FALSE) +
  #scale_size_area(max_size = 5) +
  scale_size_continuous(range = c(.2, 5)) + # Adjust size scale for better visibility
  scale_color_manual(values = color_palette) +  
  guides(color = guide_legend(override.aes = list(alpha = 1)),  
         size = guide_legend(override.aes = list(alpha = 1), title = "RPKM")) +  
  theme_bw() +
  theme(legend.position = "right",
        legend.key.width = unit(1, "cm"),
        legend.key.height = unit(0.43, "cm"),
        legend.text = element_text(size = 8)) +
  labs(x = "Longitude", y = "Latitude", color = "Gene", size = "RPKM", title = "Map of All ARG Presence") +
  coord_quickmap(xlim = c(-83.5, -82.9), ylim = c(41.6, 42.089)) +
  facet_wrap(~type_name)

# Save the map to a file
ggsave("results/All_ARGs_map.pdf", width = 12, height = 7, scale = 1) 
ggsave("results/All_ARGs_map.png", width = 12, height = 7, scale = 1) 
```


Map for Aac3 
```{r}
library(ggmap)
library(ggplot2)
library(dplyr)

register_google(maps_api_key) 

google_map <- get_googlemap(center = c(lon = -83.326801, lat = 41.778739), zoom = 10, maptype = "terrain")

ggmap(google_map) +
  geom_point(data = per_gene_rpkm_summary |> 
               filter(SampleID %in% public_metagenome_samples$SampleID) |>
               filter(gene == "Aac3"),  # Filter for Aac3 instead of ArsD
             aes(lon, lat, color = gene, size = rpkm), alpha = 0.5, inherit.aes = FALSE) +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude", color = "Gene", size = "RPKM", title = "Map of Aac3 Presence") +
  coord_quickmap(xlim = c(-83.5, -82.9), ylim = c(41.6, 42.089)) 

ggsave("results/Aac3_map.pdf", width = 12, height = 7, scale = 1) 
ggsave("results/Aac3_map.png", width = 12, height = 7, scale = 1) 

```



Number of MRG samples over time
```{r}
library(ggplot2)
library(dplyr)

selected_genes <- c("MerA", "MerP", "MerB", "MerC", "MerD", "MerE", "MerR", "MerT", "MerF", "ArsD", "PbrT", "PbrA")
selected_years <- c(2014, 2019)

# Calculate total samples per year
total_samples_per_year <- per_gene_rpkm_summary %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  filter(!is.na(year)) %>%
  filter(year %in% selected_years) %>%
  group_by(year) %>%
  select(SampleID, year) %>% distinct() %>% 
  summarise(total_samples = n(), .groups = 'drop')

# Join the totals back to the main summary
gene_year_summary <- per_gene_rpkm_summary %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  filter(rpkm > 1e-05) %>%
  filter(!is.na(year)) %>%
  filter(gene %in% selected_genes) %>%
  filter(year %in% selected_years) %>%
  select(gene, year, SampleID) %>% distinct() %>%
  group_by(gene, year) %>%
  summarise(n_samples = n(), .groups = 'drop') %>%
  left_join(total_samples_per_year, by = "year") %>%
  mutate(percentage = (n_samples / total_samples) * 100)

# Plot the graph with percentage
ggplot(gene_year_summary, aes(x = gene, y = percentage, fill = factor(year))) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  labs(x = "Gene", y = "Percentage of Samples", title = "Percentage of Samples Containing Metal Resistant Genes in 2014 and 2019", fill = "Year") +
  scale_fill_manual(values = c("2014" = "purple", "2019" = "pink")) + # Optional: Set custom colors for the bars
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) # Rotate x labels for readability

# Save the plots
ggsave("results/Percentage_of_Samples_Containing_MRGs_in_2014_and_2019.pdf", width=8, height=4, scale=1.5)
ggsave("results/Percentage_of_Samples_Containing_MRGs_in_2014_and_2019.png", width=8, height=4, scale=1.5)

```
Number of ARG Samples Over Time
```{r}
library(ggplot2)
library(dplyr)

selected_genes <- c("AmpC", "ermA", "ermC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                    "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")

# Calculate the total number of samples per year
total_samples_per_year <- per_gene_rpkm_summary %>%
  mutate(year = as.numeric(as.character(year))) %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  filter(!is.na(year)) %>%
  filter(year %in% c(2014, 2019)) %>%
  group_by(year) %>%
  select(SampleID, year) %>% distinct() %>% 
  summarise(total_samples = n(), .groups = 'drop')

# Join the totals back to the main summary and calculate the percentage
arg_year_summary <- per_gene_rpkm_summary %>%
  mutate(year = as.numeric(as.character(year))) %>%
  filter(SampleID %in% public_metagenome_samples$SampleID) %>%
  filter(rpkm > 1e-05) %>%
  filter(!is.na(year)) %>%
  filter(gene %in% selected_genes) %>%
  filter(year %in% c(2014, 2019)) %>%
  group_by(gene, year) %>%
  select(gene, year, SampleID) %>% distinct() %>%
  summarise(n_samples = n(), .groups = 'drop') %>%
  left_join(total_samples_per_year, by = "year") %>%
  mutate(percentage = (n_samples / total_samples) * 100)

# Plot the graph with percentage values
ggplot(arg_year_summary, aes(x = gene, y = percentage, fill = factor(year))) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  labs(x = "Gene", y = "Percentage of Samples", 
       title = "Percentage of Samples Containing Antibiotic Resistant Genes in 2014 and 2019", fill = "Year") +
  scale_fill_manual(values = c("2014" = "salmon", "2019" = "skyblue")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.title = element_blank()) # Rotate x labels and remove legend title

# Save the plots
ggsave("results/Percentage_of_Samples_Containing_ARGs_in_2014_and_2019.pdf", width = 10, height = 5, scale = 1.5)
ggsave("results/Percentage_of_Samples_Containing_ARGs_in_2014_and_2019.png", width = 10, height = 5, scale = 1.5)
```


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)

# Ensure year is a factor
per_gene_rpkm_summary$year <- as.factor(per_gene_rpkm_summary$year)

# Define your genes of interest
genes_of_interest <- c("AmpC", "ErmA", "ErmC", "TEM-1B", "MerX", "AmpR", "Aac3", "BcrC", "TetA", "TetB", 
                       "IpxA", "Aac(3)-Il", "VanH", "VanA", "VanX", "FosA3", "FosA", "GyrA", "GyrB")

# Filter the data for the genes of interest and the years 2014 and 2019
arg_data_filtered <- per_gene_rpkm_summary %>%
      filter(SampleID %in% public_metagenome_samples$SampleID) %>%

  filter(gene %in% genes_of_interest, year %in% c("2014", "2019"))

# Create a wide format of the data
arg_data_wide <- arg_data_filtered %>%
  pivot_wider(names_from = gene, values_from = rpkm, values_fill = list(rpkm = 0))

# Calculate the relative abundance percentages
arg_data_wide <- arg_data_wide %>%
  rowwise() %>%
  mutate(Total = sum(c_across(all_of(genes_of_interest))),
         across(all_of(genes_of_interest), ~ .x / Total * 100)) %>%
  ungroup()

# Transform the data to a long format
arg_data_long <- arg_data_wide %>%
  pivot_longer(cols = all_of(genes_of_interest), names_to = "gene", values_to = "Percentage") %>%
  filter(Percentage > 0)

# Order the genes by total percentage
gene_order <- arg_data_long %>%
  group_by(gene) %>%
  summarize(Total_Percentage = sum(Percentage), .groups = 'drop') %>%
  arrange(-Total_Percentage) %>% 
  pull(gene)

arg_data_long$gene <- factor(arg_data_long$gene, levels = rev(gene_order))

# Define gene colors (You can reuse the ones provided or create your own)
gene_colors <- c("AmpC" = "magenta", "ermA" = "limegreen", "ermC" = "blueviolet", "TEM-1B" = "orange1", 
                 "MerX" = "yellow", "AmpR" = "skyblue", "Aac3" = "grey", "BcrC" = "turquoise", 
                 "TetA" = "orangered", "TetB" = "gold", "IpxA" = "green4", "Aac(3)-Il" = "tan4",
                 "VanH" = "salmon", "VanA" = "deeppink", "VanX" = "darkorange", "FosA3" = "dodgerblue4",
                 "FosA" = "cyan3", "GyrA" = "seagreen3", "GyrB" = "sienna3")

# Create the plot
p <- ggplot(arg_data_long, aes(x = year, y = Percentage, fill = gene)) + 
  geom_bar(stat = "identity", position = "fill", width = 0.3) + 
  scale_fill_manual(values = gene_colors) +  
  theme_bw() +
  labs(x = "Year", y = "Antibiotic Resistance Gene Relative Abundance (%)", 
       title = "Relative Abundance of Antibiotic Resistant Genes Over Years") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # Rotate x labels for readability

# Print the plot
print(p)

# Save the plot
ggsave("results/ARG_relative_abundance_over_years.pdf", plot = p, width=10, height=5, scale=1.5)
ggsave("results/ARG_relative_abundance_over_years.png", plot = p, width=10, height=5, scale=1.5)

```

```{r}


# Convert the year to a factor if it's not already
per_gene_rpkm_summary$year <- as.factor(per_gene_rpkm_summary$year)


# Define genes and years of interest
genes_of_interest <- c("MerA", "MerP", "MerB", "MerC", "MerD", "MerE", "MerR", "MerT", "MerF", "ArsD", "PbrT", "PbrA")
years_of_interest <- c("2014", "2019")

# Filter the data for genes of interest and the years 2014 and 2019
arg_data_filtered <- per_gene_rpkm_summary %>%
    filter(SampleID %in% public_metagenome_samples$SampleID) %>%

  filter(gene %in% genes_of_interest, year %in% years_of_interest)


# Proceed with the code for creating a wide format, calculating percentages, and plotting

arg_data_wide <- arg_data_filtered %>%
  pivot_wider(names_from = gene, values_from = rpkm, values_fill = list(rpkm = 0))

existing_genes <- intersect(genes_of_interest, names(arg_data_wide))

arg_data_wide <- arg_data_wide %>%
  rowwise() %>%
  mutate(Total = sum(c_across(all_of(existing_genes))),
         across(all_of(existing_genes), ~ .x / Total * 100)) %>%
  ungroup()

arg_data_long <- arg_data_wide %>%
  pivot_longer(cols = all_of(existing_genes), names_to = "gene", values_to = "Percentage") %>%
  filter(Percentage > 0)

gene_order <- arg_data_long %>%
  group_by(gene) %>%
  summarize(Total_Percentage = sum(Percentage), .groups = 'drop') %>%
  arrange(-Total_Percentage) %>% 
  pull(gene)

arg_data_long$gene <- factor(arg_data_long$gene, levels = rev(gene_order))

gene_colors <- c(
  "MerA" = "#4E79A7", # Blue
  "MerP" = "#F28E2B", # Orange
  "MerB" = "#E15759", # Red
  "MerC" = "#76B7B2", # Cyan
  "MerD" = "#59A14F", # Green
  "MerE" = "#EDC948", # Yellow
  "MerR" = "#B07AA1", # Purple
  "MerT" = "#FF9DA7", # Pink
  "MerF" = "#9C755F", # Brown
  "ArsD" = "#BAB0AC", # Gray
  "PbrT" = "#D37295", # Magenta
  "PbrA" = "#A9A9A9"  # Dark Gray
)

# Create the plot
ggplot(arg_data_long, aes(x = year, y = Percentage, fill = gene)) + 
  geom_bar(stat = "identity", position = "fill", width = 0.3) +  # Adjust the width here
  scale_fill_manual(values = gene_colors) + 
  theme_bw() +
  labs(x = "Year", y = "Metal Resistance Gene Relative Abundance (%)", 
       title = "Relative Abundance of Metal Resistant Genes Over Years") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


# Saving the plots to files is the same
ggsave("results/MRG_relative_abundance_over_years.pdf", width=10, height=5, scale=1.5)
ggsave("results/MRG_relative_abundance_over_years.png", width=10, height=5, scale=1.5)

```

```{r}
if (!requireNamespace("pheatmap", quietly = TRUE)) {
    install.packages("pheatmap")
}

library(pheatmap)


data_matrix <- per_gene_rpkm_summary %>%
  dplyr::select(gene, SampleID, rpkm) %>%
  tidyr::spread(key = SampleID, value = measurement)

data_matrix <- as.matrix(data_matrix[,-1]) 
row.names(data_matrix) <- your_data$gene  

log2_data_matrix <- log2(data_matrix + 1)

pheatmap(log2_data_matrix,
         scale = "row",  
         clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean",
         clustering_method = "complete",
         color = colorRampPalette(c("navy", "white", "firebrick3"))(255),  
         border_color = NA,  
         cellwidth = 10,
         cellheight = 10,
         fontsize = 6,
         legend = TRUE)

ggsave("heatmap.pdf", width = 10, height = 8)
ggsave("heatmap.png", width = 10, height = 8, dpi = 300)

```
per_gene_rpkm_summary csv/pdf download file
```{r}
if (!requireNamespace("googledrive", quietly = TRUE)) {
    install.packages("googledrive")
}
library(googledrive)

#drive_auth()

write.csv(per_gene_rpkm_summary, "data/for_pub_supplemental/per_gene_rpkm_summary.csv", row.names = FALSE)

# file <- drive_upload(media = "per_gene_rpkm_summary.csv")
# 
# drive_share(file, role = "reader", type = "anyone")
# 
# file_id <- as_id(file)
# link <- sprintf("https://drive.google.com/file/d/%s/view", file_id)
# print(link)
```

```{r}
if (!requireNamespace("ape", quietly = TRUE)) {
    install.packages("ape")
}
if (!requireNamespace("ggtree", quietly = TRUE)) {
    install.packages("ggtree")
}

library(ape)
library(ggtree)

# Assuming the 'lineage' column of 'gene_info' contains Newick strings
# We'll use the first lineage as an example. If you have multiple, you'll need to iterate over them or specify which one to use.

# Read the Newick string into a phylogenetic tree object
tree <- read.tree(text = gene_info$lineage[1])

# Plot the tree with ggtree
p <- ggtree(tree) + 
  geom_tiplab() +  # Add tip labels
  theme_tree2()    # Use a cleaner theme suited for trees

# Visualize the tree
print(p)

# Save the tree to a file
ggsave("phylogenetic_tree.png", p, width = 10, height = 10, dpi = 300)
ggsave("phylogenetic_tree.pdf", p, width = 10, height = 10)


```
```{r}
is_newick_format <- function(string) {
  pattern <- "^\\(.*\\);$"
  grepl(pattern, string, perl = TRUE)
}

if (is_newick_format(gene_info$lineage[1])) {
  cat("The lineage information appears to be in the Newick format.\n")
} else {
  cat("The lineage information does not appear to be in the Newick format.\n")
}

```

GLAMR Datasets sheets
```{r}
# Install and load the required packages
if (!requireNamespace("googledrive", quietly = TRUE)) {
    install.packages("googledrive")
}
if (!requireNamespace("googlesheets4", quietly = TRUE)) {
    install.packages("googlesheets4")
}
library(googledrive)
library(googlesheets4)

# Authenticate Google account for both googledrive and googlesheets4
drive_auth()
gs4_auth()

# Save your data frame to a Google Sheet
GLAMR_datasets <- read.csv("GLAMR_sample_data.csv")

# Create a new Google Sheet and upload the data frame
sheet <- gs4_create("GLAMR Datasets", sheets = list("Sheet1" = GLAMR_datasets))

# Print the link to the Google Sheet
sheet_url <- sheet$spreadsheet_url
print(sheet_url)



```



```{r}
write.csv(filtered_samples,"filtered_samples.csv")
```

```{r}
write.csv(public_glmr_datasets,"public_GLAMR_dataset.csv")

```



Filtered Samples Sheets
```{r}
# Install and load the required packages
if (!requireNamespace("googledrive", quietly = TRUE)) {
    install.packages("googledrive")
}
if (!requireNamespace("googlesheets4", quietly = TRUE)) {
    install.packages("googlesheets4")
}
library(googledrive)
library(googlesheets4)

# Authenticate Google account for both googledrive and googlesheets4
drive_auth()
gs4_auth()

# Save your data frame to a Google Sheet
filtered_samples <- read.csv("filtered_samples.csv")

# Create a new Google Sheet and upload the data frame
sheet <- gs4_create("Filtered Samples", sheets = list("Sheet1" = filtered_samples))

# Print the link to the Google Sheet
sheet_url <- sheet$spreadsheet_url
print(sheet_url)

```


```{r}
# Get a data frame of all installed packages and their versions
installed_packages <- as.data.frame(installed.packages()[, c("Package", "Version")])

# Display the data frame
print(installed_packages)

```

